Spin canonical rings of log stacky curves

### 2015-07-02 stacky_write_up/spin_cring.tex
Edits to sections 8 and 9

This paper leaves open many directions for further research.

\begin{enumerate}
	\item One direction for further research is to 
		extend the results of this paper to divisors $D \in \di \sx$ on a 
		stacky curve $\sx,$ where $nD \sim K.$ The canonical rings of such 
		divisors would translate to rings of fractional weight modular 
		forms. As defined in Adler and Ramanan (in the passage immediately 
		following \cite[Corollary 24.5]{adler:moduli}), a {\bf modular form 
		of fractional weight} $\frac{2a}{n}$ on $X(p)$ is a section of a 
		line bundle $\eta$ so that there exists a line bundle $\gamma$ with 
		$\gamma^{\otimes a} \cong \eta$ and $\gamma^{n} \cong \Omega_{X(p)}^
		1$. Adler and Ramanan \cite[Appendix 2]{adler:moduli} construct an 
		explicit basis of weight $\frac{4}{5}$ modular forms for $X(11)$. 
		Fractional modular forms also make an appearance in Milnor \cite[$
		\mathsection$ 6]{milnor:fractional-weight}. In this work, Milnor 
		shows that for a triangle group $\Gamma$ of signature $(0; e_1,e_2,e
		_3; 0)$ there exist generators $f_1,f_2,f_3$ of a certain ring of 
		fractional weight modular forms, satisfying $f_1^{e_1}+f_2^{e_2}+f_3
		^{e_3} = 0$.

		Having motivated the importance of fractional weight modular forms, we pose the following question:

		\begin{question}
		\label{ques:fractional-weight}
			If $\sx$ is a stacky curve and $D \in \di \sx$ with $nD \sim K,$ 
			where $K$ is the canonical divisor of $\sx$, can one bound the 
			degrees of generators and relations of $R_D$?
		\end{question}
		Note that that the above question is answered affirmatively in the 
		case that $\sx$ has genus $0$ and $D$ is effective, as follows by 
		inductively applying Lemma ~\ref{lem:sat-1}: If $\sx$ has signature 
		$(g;e_1, \ldots, e_r; \delta)$ then $D$ is generated in degree at 
		most $e = \max(e_1, \ldots, e_r)$ with relations in degree at most
		$2e$. Additionally, it may be possible to modify the proof of Lemma
		~\ref{lem:raise-stacky-order} to extend to the setting of fractional 
		weight modular forms. Given suitable generalizations of these 
		lemmas, one may be able to approach the problem of finding 
		generators and relations for fractional weight modular forms using 
		techniques similar to those in this paper.

	\item In Voight and Zureick Brown \cite[Definition 2.2.7]{vzb:stacky},
		the notion of a generic initial ideal is defined, which 
		encapsulates the idea of whether the relations for $R_\halfcan$ are 
		generically chosen, in a precise sense. The proof of Theorem
		~\ref{thm:main} decidedly does not yield a generic initial ideal, 
		as generators and relations are constructed in a highly nongeneric 
		manner. In particular, Lemma ~\ref{lem:raise-stacky-order} 
		constructs generators with non-maximal pole orders at certain 
		points, making the relations non-generic. 
		\begin{question}
		\label{ques:generic-initial}
			If $(\sx,\Delta,L)$ is a log spin curve with signature $\sigma$,
			does there exist a function $\sigma \mapsto d(\sigma)$ so that
			the generic initial ideal {\rm(} with some specified ordering{\rm)}
			is generated in degree at most $d(\sigma)$? Can one write down the
			generic initial ideal explicitly?
		\end{question}

	\item As noted in Remark ~\ref{rem:explicit-generators}, the 
		proof of Theorem ~\ref{thm:main} given in this paper gives an 
		explicit procedure for computing the generators and relations of $R_
		L$ when the genus of $\sx$ is $0$ or $1$.
		\begin{question}
		\label{ques:algorithm}
			Can one give an algorithm for explicitly computing the generators 
			and relations for an arbitrary genus $g \geq 2$ curve?
		\end{question}
		Suppose the generators and relations of $R_{\halfcan_X}$ are known, 
		where $(\sx, \Delta, L)$ is a spin canonical ring, $X$ is the 
		coarse space of $\sx$, and $L_X \leq L$ is a spin canonical divisor 
		of $X$. In this case our proof of Theorem ~\ref{thm:main} yields an 
		algorithm for determining the generators of $L$. This raises the 
		following Petri-like question:
		\begin{question}
		\label{ques:spin-canonical-petri}
			Given a {\rm(}non-stacky {\rm)} curve $X$ and a spin canonical 
			divisor $L \in \di X$, can one write down a set of minimal 
			generators and minimal relations for $R_\halfcan$? If so, can one give the 
			generic ideal of $R_\halfcan$ or the generic initial ideal of $R_\halfcan$, as 
			defined in Voight and Zureick Brown \cite[Definition 2.2.7]
			{vzb:stacky}?
		\end{question}
		In particular, if Question ~\ref{ques:spin-canonical-petri} were 
		answered affirmatively one would be able to write down explicit 
		generators and relations for $R_\halfcan$ on a stacky curve, essentially 
		by inductively applying Lemma ~\ref{lem:raise-stacky-order}.
\todo{Aaron, I tried to edit the following question, but it still seems a bit imprecise, and the question in particular is a bit of a long sentence.}
	\item In Subsection \ref{ssec:bounds-high-genus}, we reference 
		the work of Reid \cite[Theorem 3.4]{reid:infinitesimal}. We use his 
		proof of that the 
		spin canonical ring is generated in degree at most 5 with
		relations in degree at most 10 in the 			
		non-log, non-stacky case when genus is at least 2. We extend this 
		bound of 5 and 10 to the log case, and then apply our inductive 
		lemmas to add stacky points and obtain bounds of $e = \max(5,e_1, \ldots, e_r)$ and $2e$.  Reid \cite[Theorem 3.4]{reid:infinitesimal} in fact proves something slightly stronger: that in 
		most cases his bound is actually 3 and 6 with well-characterized 
		exceptions.
		\begin{question}\label{ques:reduce-5,10-to-4,8}
			In the case that the genus $g \geq 2,$ can we reduce the bounds in Theorem \ref{thm:main} on 
			the degrees of generation and relations from $e = \max(5,e_1,\ldots, e_r)$ and $\max(10, 2e_1, \ldots, 2e_r)$ to $\max(4,e_1, \ldots, e_r)$ and $
			\max(8, 2e_1, \ldots, 2e_r),$ apart from a finite list of exceptions?
		\end{question}
		\begin{rem}\label{rem:reduce-3,6}
			It is unlikely that this can be reduced further for the case 
			of stacky curves, because Lemma \ref{lem:sat-2}, which allows 
			us to add a stacky point in genus $g > 0$ saturated at 2, 
			requires a generator in degree 4 and potentially a relation in 
			degree 8, even when inductively adding on stacky points of 
			stabilizer order 3. However, perhaps in the non-stacky case, one 
			may be able able to generalize the result of Reid \cite[Theorem 3.4
			]{reid:infinitesimal} to show that log spin canonical rings can be 
			generated in degree at most 3 with relations in degree at most 6, 
			apart from a well characterized list of exceptions.
			\end{rem}
	\item While Theorem ~\ref{thm:main} gives a minimal set of 
		generators for the log spin canonical ring $R_\halfcan,$ the set of 
		relations is not minimal. In many of the $g = 0$ and $g=1$ cases, 
		it is not too difficult to see that our inductive procedure also 
		yields a minimal set of relations for $R_\halfcan$. One might investigate 
		whether the generators and relations given by the inductive proof 
		of Theorem ~\ref{thm:main} are always minimal.
		\end{enumerate}

@@

\section{Acknowledgments}
We are grateful to David Zureick-Brown for introducing us to this
field of study, providing incredibly helpful guidance, and being an
excellent project mentor. We also thank Ken Ono and the Emory 
University Number Theory REU for arranging our project and 
providing a great environment for mathematical learning and 
collaboration. We thank Jorge Sentieiro Neves for pointing out the
bound given by Reid's work on the maximal degrees of minimal
generators and relations of spin canonical rings in high genera.
Finally, we gratefully acknowledge that our research was
financially supported by NSF Grant Award Number 1250467 via the
Emory University Number Theory REU. We deeply appreciate all of the
support that has made our work possible.

>>>>

This paper leaves open many directions for further research.\todo{Peter: This sentence sounds weird.  It's also kind of not true; as David said, we are kind of closing the topic}

\begin{enumerate}
	\item One direction for further research is to 
		extend the results of this paper to divisors $D \in \di \sx$ on a 
		stacky curve $\sx,$ where $nD \sim K.$ The canonical rings of such 
		divisors translate to rings of fractional weight ($\frac{2}{n}$) modular 
		forms. As defined in Adler and Ramanan (in the passage immediately 
		following \cite[Corollary 24.5]{adler:moduli}), a {\bf modular form 
		of fractional weight} $\frac{2a}{n}$ on $X(p)$ is a section of a 
		line bundle $\eta$ so that there exists a line bundle $\gamma$ with 
		$\gamma^{\otimes a} \cong \eta$ and $\gamma^{n} \cong \Omega_{X(p)}^
		1$. Adler and Ramanan \cite[Appendix 2]{adler:moduli} construct an 
		explicit basis of weight $\frac{4}{5}$ modular forms for $X(11)$. 
		Fractional modular forms also make an appearance in Milnor \cite[$
		\mathsection$ 6]{milnor:fractional-weight}. In this work, Milnor 
		shows that for a triangle group $\Gamma$ of signature $(0; e_1,e_2,e
		_3; 0)$ there exist generators $f_1,f_2,f_3$ of a certain ring of 
		fractional weight modular forms, satisfying $f_1^{e_1}+f_2^{e_2}+f_3
		^{e_3} = 0$.  \todo{Peter: I don't know if it is so worth while giving an explicit definition and precisely defining the work already done.  A general outline might suffice and be easier/shorter to read.}

		This motivates the following question:

\todo{Peter: Would this be better as a conjecture?}
		\begin{question}
		\label{ques:fractional-weight}
			If $\sx$ is a stacky curve and $D \in \di \sx$ with $nD \sim K$, 
			where $K$ is the canonical divisor of $\sx$, can one bound the 
			degrees of generators and relations of $R_D$?
		\end{question}
		When $g=0$ and $D$ is effective, inductively applying Lemma ~\ref{lem:sat-1}  
		gives an affirmative answer to this question: If $\sx$ has signature 
		$(g;e_1, \ldots, e_r; \delta)$ then $D$ is generated in degree at 
		most $e = \max(e_1, \ldots, e_r)$ with relations in degree at most
		$2e$. 
		It may be possible to modify the proof of Lemma
		~\ref{lem:raise-stacky-order} to extend to the setting of fractional 
		weight modular forms. 
		Suitable generalizations of the 
		lemmas of Section ~\ref{sec:induction} might allow one to 
		follow similar questions to this paper and provide a general 
		answer to Question ~\ref{ques:fractional-weight}.

	\item Voight and Zureick Brown define
		the notion of a generic initial ideal \cite[Definition 2.2.7]{vzb:stacky}, which 
		encapsulates the idea of whether the relations for $R_\halfcan$ are 
		generically chosen, in a precise sense. 
		The proof of Theorem
		~\ref{thm:main} is decidedly non-generic. 
		In particular, Lemma ~\ref{lem:raise-stacky-order} 
		constructs generators with non-maximal pole orders at certain 
		points, making the relations non-generic. 
		\begin{question}
		\label{ques:generic-initial}
			If $(\sx,\Delta,L)$ is a log spin curve with signature $\sigma$,
			does there exist a function $\sigma \mapsto d(\sigma)$ so that
			the generic initial ideal {\rm(}with some specified ordering{\rm)}
			is generated in degree at most $d(\sigma)$? Can one write down the
			generic initial ideal explicitly?
		\end{question}\todo{Peter: This whole question is kind of confusing.  I don't know what it is trying to say.}

	\item As noted in Remark ~\ref{rem:explicit-generators}, the 
		proof of Theorem ~\ref{thm:main} 
		%given in this paper 
		gives an 
		explicit procedure for computing the generators and relations of $R_
		\halfcan$ when the genus of $\sx$ \todo{Peter: Minor note: should we say genus of $\sx$ or genus of $X$; did we actually define that a genus of a stacky curve equals that of its coarse space?} is $0$ or $1$.  \todo{Peter: This question also seems like it might be the most important (as far as modular forms applications go.  Maybe we should put it first?}
		\begin{question}
		\label{ques:algorithm}
			Is there an explicit procedure for computing the generators 
			and relations for an arbitrary genus $g \geq 2$ curve?
		\end{question}
		Suppose the generators and relations of $R_{\halfcan_X}$ are known, 
		where $(\sx, \Delta, L)$ is a spin canonical ring, $X$ is the 
		coarse space of $\sx$, and $L_X \leq L$ is a spin canonical divisor 
		of $X$. In this case our proof of Theorem ~\ref{thm:main} yields an 
		algorithm for inductively determining the generators of $L$. 
		This reduces Question \ref{ques:algorithm} to the 
		following Petri-like question:
		\begin{question}
		\label{ques:spin-canonical-petri}
			Given a {\rm(}non-stacky {\rm)} curve $X$ and a spin canonical 
			divisor $L \in \di X$, can one write down a set of minimal 
			generators and minimal relations for $R_\halfcan$? If so, can one give the 
			generic ideal of $R_\halfcan$ or the generic initial ideal of $R_\halfcan$, as 
			defined in Voight and Zureick Brown \cite[Definition 2.2.7]
			{vzb:stacky}?
		\end{question}
		In particular, if Question ~\ref{ques:spin-canonical-petri} were 
		answered affirmatively one would be able to write down explicit 
		generators and relations for $R_\halfcan$ on a stacky curve, essentially 
		by inductively applying Lemma ~\ref{lem:raise-stacky-order}.
\todo{Aaron, I tried to edit the following question, but it still seems a bit imprecise, and the question in particular is a bit of a long sentence.}
\todo{Peter @ Aaron: I think the last sentence of the question can be removed; generic isn't relevant in this part of further questions (it was relevant in part (2)).  Also the sentence after the question can be removed.}
	\item In Subsection \ref{ssec:bounds-high-genus}, we reference 
		the work of Reid \cite[Theorem 3.4]{reid:infinitesimal}. We use his 
		proof that the 
		spin canonical ring is generated in degree at most 5 with
		relations in degree at most 10 in the 			
		non-log, non-stacky case when genus is at least 2. We extend this 
		bound of 5 and 10 to the log case, and then apply our inductive 
		lemmas to add stacky points and obtain bounds of $e = \max(5,e_1, \ldots, e_r)$ and $2e$.  However, Reid in fact proves something slightly stronger \cite[Theorem 3.4]{reid:infinitesimal}: that in 
		most cases his bound is actually 3 and 6 with well-characterized 
		exceptions.
		Generalizing this slightly stronger bound to the (non-stacky) log case case would allow us to inductively apply the lemmas from Section ~\ref{sec:induction} and improve Theorem ~\ref{thm:main} as follows:
		\begin{conjecture}\label{ques:reduce-5,10-to-4,8}
			When $g \geq 2,$ the bounds in Theorem \ref{thm:main} on 
			the degrees of generation and relations can be reduced from 
			$e = \max(5,e_1,\ldots, e_r)$ and $\max(10, 2e_1, \ldots, 2e_r)$ to 
			$e'=\max(4,e_1, \ldots, e_r)$ and $2e'$, 
			apart from a finite list of exceptions?
		\end{conjecture}
		\todo{Peter: I don't like a finite list of exceptions.  The list contains infinitely many non-isomorphic stacky curves.  Finite family might also be imperfect.  Can we say except when $g=2$ (I believe there will be infinitely many counterexamples there) or it lies in a finite list of exception?  Do we believe that to be true?}
		\begin{rem}\label{rem:reduce-3,6}
			It is unlikely that these bounds can be reduced further to 
			$e''=\max(3, e_1, \ldots, e_r)$ and $2e''$, 
			because Lemma \ref{lem:sat-2}'s construction always
			requires a generator in degree 4 and potentially a relation in 
			degree 8, when adding a stacky point with stabilizer order 3. 
%			Perhaps in the non-stacky case, one 
%			may be able able to generalize the result of Reid \cite[Theorem 3.4
%			]{reid:infinitesimal} to show that log spin canonical rings can be 
%			generated in degree at most 3 with relations in degree at most 6, 
%			apart from a well characterized list of exceptions.
			\end{rem}
	\item While Theorem ~\ref{thm:main} gives a minimal set of 
		generators \todo{Peter: I believe this is actually not true.  If $R_{\halfcan'}$ is minimally generated by $x_1, \ldots x_m$ and $R_\halfcan$ is generated by $y_1, \ldots, y_n$ over $R_{\halfcan'}$, then $R_{\halfcan}$ is {\emph{generated}} by $x_1, \ldots, x_m, y_1, \ldots y_n$ but not necessarily minimally.  It's possible to have something like $y_1y_3-y_2^2=x_i$.}
		for the log spin canonical ring $R_\halfcan,$ the set of 
		relations is not minimal. In many of the $g = 0$ and $g=1$ cases, 
		it is not too difficult to see that our inductive procedure also 
		yields a minimal set of relations for $R_\halfcan$. One might investigate 
		whether the generators and relations given by the inductive proof 
		of Theorem ~\ref{thm:main} are always minimal.
		\end{enumerate}

@@

\section{Acknowledgments}
We are grateful to David Zureick-Brown for introducing us to the study of stacky canonical rings, providing invaluable guidance, and being \todo{Peter: using passive voice (i.e. ``being'') makes this sound kind of generic} an
excellent project mentor. We also thank Ken Ono and the Emory 
University Number Theory REU for arranging our project and 
providing a great environment for mathematical learning and 
collaboration. \todo{Peter: Does it make sense to thank the REU? Ken essentially is the REU.} We thank Jorge Sentieiro Neves for pointing out the
bounds given by Reid's work on the maximal degrees of minimal
generators and relations of spin canonical rings in high genera.
Finally, we gratefully acknowledge the financial support given by NSF Grant Award Number 1250467 via the
Emory University Number Theory REU. 
We deeply appreciate all of the
support that has made our work possible.
